### 2020-12-16 MetaWebQSSAnalysis.Rmd
First commit

---
title: "From metawebs to regional and local"
author: "L.A.S."
output: html_document
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup, eval=TRUE,echo=FALSE }
load(".RData")
# rm(list=lsf.str())
# rm(list=ls())
# oldcd <-getwd()
require(ggplot2)
require(igraph)
require(dplyr)
require(stringr)
require(devtools)
require(meweasmo)
#  install_github("lsaravia/meweasmo")
require(multiweb)
#  install_github("lsaravia/multiweb")

source("R/network_fun.r")

```

# Build network without bird species from aves data.frame

```{r buildNetworkNoBirds, eval=F,echo=FALSE,message=T,warning=T}

#
# Read Networks 
#
#
meta <- readNetwork("Data/Meta.dat",edgeListFormat = 2)

# Meta es la simplificacion de meta1 
#meta <- simplify(meta, remove.multiple = TRUE, remove.loops = FALSE)

wedd <- readNetwork("Data/Wedd.dat",edgeListFormat = 2)

# Check that local webs are included in the metaweb
#
ecount(difference(wedd,meta)) == 0 
length(setdiff( V(wedd)$name,V(meta)$name) ) ==0
basal_species(wedd)                             #### SOME BASAL SPECIES WRONG IN OLD WEDDELL!!!!!!!!

# Vertex that are Birds
#
# Extract the birds from the network
#
require(taxize)
dfn <- igraph::as_data_frame( wedd, what="vertices" )
can_names <- gnr_resolve(dfn$name, best_match_only = TRUE, canonical=TRUE)
class_names <- tax_name(query =can_names$matched_name2  , get = "class", db = "ncbi")
aves <- class_names %>% filter( class =="Aves")

siBird <- V(wedd)$name %in% aves$query
sum(siBird) == nrow(aves)   # ALl rows in aves must be the in wedd igraph object  
noBird <- delete_vertices(wedd, siBird)
rm(dfn,nod,g,n0,g1,lnk,siBird,lred,can_names,class_names)
rm(list=names(Filter(is.function, mget(ls(all=T)))))
saveRDS(noBird,"Data/noBird_ig.rds")
save.image()

# degree 
V(noBird)[ degree(noBird) == 0]
siBird <- V(wedd)$name %in% aves$query

bsp <- tibble::rownames_to_column(as.data.frame(degree(wedd, v=siBird, mode=c("out")))) 
names(bsp) <- c("Species","out_degree")
bsp1 <- tibble::rownames_to_column(as.data.frame(degree(wedd, v=siBird, mode=c("in")))) 
names(bsp1) <- c("Species","in_degree")
bsp <- inner_join(bsp1,bsp)
knitr::kable(bsp %>% arrange(desc(in_degree)))

```

# Calc Metrics/QSS for empirical networks

```{r calcMetrics, eval=F,echo=F,message=T,warning=T}

ecount(wedd)
ecount(noBird)
require(tictoc)
tic()
qss <- bind_cols(calc_topological_indices(wedd), calc_QSS(wedd,10000,48)) %>% mutate(Network="Weddell")
toc()

qss <- bind_rows(qss, bind_cols(calc_topological_indices(noBird), calc_QSS(noBird,10000,48)) %>% mutate(Network="Nobirds"))

# QSS with simetric coefficients like Old simulations

qss <- qss %>% select(Network, everything())

qss1 <- calcIncoherence(wedd)  %>% mutate(Network="Weddell")
qss1 <- bind_rows(qss1, calcIncoherence(noBird) %>% mutate(Network="Nobirds"))
qss <- inner_join(qss,qss1)
rm(qss1)
save.image()

```

# Calculate Small-World-Ness

The random networks with the same number of links and species than Weddell Sea generate networks with many disconected components (15551) 
in consequence is not possible to calculate small-world-ness because the clustering coeficient is 0

```{r calcSWNess, eval=F,echo=T,message=T,warning=T}

require(future.apply)
require(tictoc)
plan(multiprocess)
ss <- vcount(wedd)
ll <- ecount(wedd)
nsim <- 1000
tic("Generate Erdos Renyi networks")
ern <- future_lapply(1:nsim, function (x) {
    e <- sample_gnm(ss, ll,directed = TRUE)
    return(e) }
    )
toc()

# calculate the number of components

ncomp <- sapply(ern, function(g) components(g)$no)
mean(ncomp)

tic("SWness for Weddell")
tp0 <- calc_swness_zscore(wedd,ern,ncores=48)
tp1 <- mutate(tp0$da, Network="Weddell")
toc()

# NoBirds Weddell
#
tic("Generate Erdos Renyi NoBirds")

ss <- vcount(noBird)
ll <- ecount(noBird)
ern <- future_lapply(1:nsim, function (x) {
    e <- erdos.renyi.game(ss, ll, type="gnm",directed = TRUE)
    return(e) }
    )
toc()

tic("SWness for NoBirds")
tp0 <- calc_swness_zscore(noBird,ern,ncores=48)
tp1 <- bind_rows(tp1,mutate(tp0$da, Network="Nobirds"))
toc()

qss <- inner_join(qss,tp1)

rm(ncomp,ll,ss,nsim,tp1,tp0,ern)
save.image()

```

# Compare Networks using curve ball randomizations

```{r compTopologicalMetrics, eval=F,echo=F,message=T,warning=T}

lred <- curve_ball(wedd,1000)
curveBallNets <- calcTopologicalIndices(lred,ncores=48)
curveBallNets <- bind_cols(curveBallNets, calcIncoherence(lred,ncores = 48)) %>% mutate(Network="Weddell")

tp0 <- calcModularitySWnessZScore(wedd,lred,ncores=48)
tp <- mutate(tp0$da, Network="Weddell")
#websTbl <- websTbl %>% select(Size:QSS)
curveBallSumm <- tp
cb1 <- tp0$sims %>% dplyr::select(modularity)
curveBallNets <- bind_cols(curveBallNets, cb1)

rm(lred,tp0,tp,cb1)
save.image()
```

```{r compTopologicalMetricsNoBird, eval=F,echo=F,message=T,warning=T}

# Add the simulations for Weddell with no birds 
#
lred <- curve_ball(noBird,1000)
tp <- calcTopologicalIndices(lred,ncores=48)
tp <- bind_cols(tp, calcIncoherence(lred,ncores = 48)) %>% mutate(Network="Nobirds")

curveBallNets <- bind_rows(curveBallNets,tp)

tp0 <- cacModularitySWnessZScore(noBird,lred,ncores=48)
tp <- mutate(tp0$da, Network="Nobirds")
#websTbl <- websTbl %>% select(Size:QSS)
curveBallSumm <- bind_rows(curveBallSumm,tp)

cb1 <- tp0$sims %>% dplyr::select(modularity)
curveBallNets <- bind_cols(curveBallNets, cb1)

rm(lred,tp0,tp,cb1)
save.image()

```

## Plot metrics interval confidence using curveBall

* ADD PLOT Bivariate confidence ellipses for metrics in function of S

```{r plotMetricsCurbeBall, eval=FALSE,echo=F,message=T,warning=T}

# Density plots
#
g1 <- ggplot(curveBallNets, aes(x=Q, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=Q,  colour=Network), linetype="dashed", size=.5) + 
    scale_color_viridis_d(labels=c("NO birds","YES Birds")) + 
    scale_fill_viridis_d(labels=c("NO birds","YES Birds"))+
    coord_cartesian(xlim=c(.69,0.75)) + xlab("Trophic Coherence")
g1

g2 <- ggplot(curveBallNets, aes(x=TLmean, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=TLmean,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d() +
    coord_cartesian(xlim=c(2.55,2.68)) + xlab("Mean Trophic Level")
g2

g3 <- ggplot(curveBallNets, aes(x=PathLength, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=PathLength,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d() #+ coord_cartesian(xlim=c(2.55,2.68)) 
g3

g4 <- ggplot(curveBallNets, aes(x=Clustering, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=Clustering,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d() 
g4

g5 <- ggplot(curveBallNets, aes(x=Omnivory, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=Omnivory,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d() 
g5

g6 <- ggplot(curveBallNets, aes(x=modularity, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=curveBallSumm, aes(xintercept=Modularity,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d() 
g6

# Deleted PathLength and Clustering
#
#
require(cowplot)
gg <- plot_grid(g2 + theme(legend.position="none"),
                g1 + theme(legend.position="none"),
                g5 + theme(legend.position="none"),
                g6+ theme(legend.position="none"),align = "h",ncol=2)
gg

legend_b <- get_legend(
  g1 + guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

# add the legend underneath the row we made earlier. Give it 10%
# of the height of one plot (via rel_heights).
gg <- plot_grid(gg, legend_b, ncol = 1, rel_heights = c(1, .1))
gg

save_plot("Figures/curveBall_metrics.png",gg,base_width=8,base_height=5,dpi=600)

#
# Anderson Darling test of distributions and effect size  
#
require(kSamples)
require(tidyr)
require(purrr)

df <- curveBallNets %>% dplyr::select(Q,TLmean,Omnivory,PathLength,Clustering,modularity) %>% map(~ad.test(. ~ Network, data=curveBallNets)) %>% map_df(~.$ad[1,])
knitr::kable(df)

df <- curveBallNets %>% group_by(Network) %>% dplyr::select(Q,TLmean,Omnivory,PathLength,Clustering,modularity) %>% summarize_if(is.numeric, median)  %>% gather("metric","medianDif" ,-Network) %>% group_by(metric) %>% summarize_if(is.numeric,~diff(.))

df1 <- curveBallNets %>% dplyr::select(Network,Q,TLmean,TLmax,Omnivory,PathLength,Clustering,modularity) %>% gather("metric","value" ,-Network) %>%   group_by(metric) %>% summarize(sdM = sd(value))

df2 <- curveBallNets %>% group_by(Network) %>% dplyr::select(Q,TLmean,Omnivory,PathLength,Clustering,modularity) %>% summarize_if(is.numeric, median) %>%  gather("metric","median" ,-Network) %>%  spread(Network,median)

df <- inner_join(df,df1)  %>% inner_join(df2) %>% mutate(effectSize=medianDif/sdM) %>% arrange(effectSize)

knitr::kable(df)

```

# Simulate a grid of parameters

```{r fitMetaWebPC, eval=F,echo=F,message=T,warning=T}

require(MetaWebAssemblyModels)

A <- get.adjacency(meta,sparse=F)
dim(A)==919

# Simulate a grid of parameters to select the best options
#
#

# Combine parameters
#
simMetaWebAssembly <- simul_metaWebAssembly(0.01,0.1,0.001, 0.01,0.1, 0.001, 10, 500, A)
simMetaWebAssembly <- bind_rows(simMetaWebAssembly, simul_metaWebAssembly(0.001,0.01,0.001, 0.1,0.9, 0.001, 10, 500, A))

saveRDS(simMetaWebAssembly,"Data/simulations_metaWebAssembly.rds")
simMetaWebAssembly <- readRDS("Data/simulations_metaWebAssembly.rds")

require(RColorBrewer)
col <- brewer.pal(8,"Dark2")

ggplot(simMetaWebAssembly,aes(S, L))+ geom_point(alpha=0.5,shape = 21,fill="white",color=col[1],size=1) +  geom_point(data=qss,aes(Size, Links,colour=Network),shape = 21, size=1) + theme_bw() + scale_color_viridis_d()
ggsave("Figures/MetaWebMdls_vs_data.png",width=6,height=6,units="in",dpi=600)
#
# Selecting simulations with 30% tolerance for both Species and Links
#

sel_tolerance_local_web(list(noBird,wedd),meta,0.30,simMetaWebAssembly)

#        m      a     S      L      C  cost alpha     n
#    <dbl>  <dbl> <dbl>  <dbl>  <dbl> <dbl> <dbl> <int>
# NoBi 0.0728 0.0241  613. 11167. 0.0298 0.583  3.01    54
# Wedd 0.0703 0.0221  626. 11546. 0.0295 0.557  3.19   143

################################################################################# ACA!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
#
# Simulations with the fitted parameters
#
tf <- 500
mig <- rep(0.0728,nrow(A))
ext <- rep(0.0241,nrow(A))
AA <- metaWebNetAssembly(A,mig,ext,tf)
plot_NetAssemblyModel(AA,400)
plot_NetAssemblyModel_eqw(AA,100)
tf <- 500
mig <- rep(0.0703,nrow(A))
ext <- rep(0.0221,nrow(A))
AA <- metaWebNetAssembly(A,mig,ext,tf)
plot_NetAssemblyModel(AA,400)
plot_NetAssemblyModel_eqw(AA,100)

```

# Simulate with fitted parameters to get species variability

```{r fittedSimulationsMetaWebAssembly, eval=F,echo=F,message=T,warning=T}
require(future.apply)
require(MetaWebAssemblyModels)
plan(multiprocess)
#plan(sequential)

A <- get.adjacency(meta,sparse=F)
dim(A)==919

# Simulate Nobirds
#
nsim <- 1000
tf <- 500
mig <- rep(0.0728,nrow(A))
ext <- rep(0.0241,nrow(A))
tini <- tf-99                  # Ultimos 100 pasos temporales

df <- future_lapply(seq_len(nsim), function(i){
      # Nobirds
      AA <- metaWebNetAssembly(A,mig,ext,tf)
      g <- graph_from_adjacency_matrix( AA$A, mode  = "directed")
      # Select only a connected subgraph graph 
      dg <- components(g)
      g <- induced_subgraph(g, which(dg$membership == which.max(dg$csize)))

      sti <- calcTopologicalIndices(g) 
      sti <- bind_cols(sti,calcIncoherence(g))
      sti <- bind_cols(sti,calc_modularity(g))
      sti <- bind_cols(sti,tibble(sdS=sd(AA$S[tini:tf]),meanS=mean(AA$S[tini:tf]),cvS=sdS/meanS))
  })

fittedMetaWebAssemblySim <- do.call(rbind,df) %>% mutate(Network="Nobirds")

# Complete Weddell
#
mig <- rep(0.0703,nrow(A))
ext <- rep(0.0221,nrow(A))

df <- future_lapply(seq_len(nsim), function(i){
      # Nobirds
      AA <- metaWebNetAssembly(A,mig,ext,tf)
      g <- graph_from_adjacency_matrix( AA$A, mode  = "directed")
      # Select only a connected subgraph graph 
      dg <- components(g)
      g <- induced_subgraph(g, which(dg$membership == which.max(dg$csize)))

      sti <- calcTopologicalIndices(g) 
      sti <- bind_cols(sti,calcIncoherence(g))
      sti <- bind_cols(sti,calc_modularity(g))
      sti <- bind_cols(sti,tibble(sdS=sd(AA$S[tini:tf]),meanS=mean(AA$S[tini:tf]),cvS=sdS/meanS))
  })

fittedMetaWebAssemblySim <- bind_rows(fittedMetaWebAssemblySim, do.call(rbind,df) %>% mutate(Network="Weddell"))

rm(df,mig,ext,A,AA,tf,tini,lred)
save.image()

```

# Plot Metaweb assembly simulations

```{r plotSimulationsMetaWebAssembly, eval=F,echo=T,message=T,warning=T}

names(fittedMetaWebAssemblySim)
names(qss)
# ggplot(fittedMetaWebAssemblySim, aes(x=Q, fill=Network)) +
#     geom_histogram( alpha=.5, position="identity") + theme_bw() + 
#     geom_vline(data=qss, aes(xintercept=Q,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d()

g1 <- ggplot(fittedMetaWebAssemblySim, aes(x=Q, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=Q,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d()

g2 <- ggplot(fittedMetaWebAssemblySim, aes(x=TLmean, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=TLmean,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d()

g3 <- ggplot(fittedMetaWebAssemblySim, aes(x=PathLength, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=PathLength,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d()

g4 <- ggplot(fittedMetaWebAssemblySim, aes(x=Clustering, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=Clustering,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d()

g5 <- ggplot(fittedMetaWebAssemblySim, aes(x=Omnivory, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=qss, aes(xintercept=Omnivory,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d()

g6 <- ggplot(fittedMetaWebAssemblySim, aes(x=Modularity, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=curveBallSumm, aes(xintercept=Modularity,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d()

require(cowplot)
gg <- plot_grid(g2 + theme(legend.position="none"),
                g3 + theme(legend.position="none"),
                g4 + theme(legend.position="none"),
                g1 + theme(legend.position="none"),
                g5 + theme(legend.position="none"),
                g6+ theme(legend.position="none"),align = "h",ncol=3)
gg

legend_b <- get_legend(
  g1 + 
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

# add the legend underneath the row we made earlier. Give it 10%
# of the height of one plot (via rel_heights).
gg <- plot_grid(gg, legend_b, ncol = 1, rel_heights = c(1, .1))

save_plot("Figures/metaWebAssem_metrics.png",gg,base_width=8,base_height=5,dpi=600)

sfit <- fittedMetaWebAssemblySim %>% group_by(Network) %>% summarize(cvS = median(cvS))

ggplot(fittedMetaWebAssemblySim, aes(x=cvS, fill=Network)) +
    geom_density( alpha=.5) + theme_bw() + 
    geom_vline(data=sfit, aes(xintercept=cvS,  colour=Network), linetype="dashed", size=.5) + scale_color_viridis_d() + scale_fill_viridis_d() + xlab("Temporal CV") 
ggsave("Figures/TempCV_MetaWebAss.png",width=6,height=6,units="in",dpi=600)

require(kSamples)
require(tidyr)
require(purrr)

df <- fittedMetaWebAssemblySim %>% dplyr::select(Q,TLmean,Omnivory,PathLength,Clustering,Modularity,cvS) %>% map(~ad.test(. ~ Network, data=fittedMetaWebAssemblySim)) %>% map_df(~.$ad[1,])
knitr::kable(df)

```

# Calculate topological roles

```{r calcTopoRoles, eval=T,echo=T,message=T,warning=T}

tic("Topo Roles")
topoR <- multiweb::calc_topological_roles(wedd,nsim=100,ncores = 48) %>% mutate(Network="Weddell")
toc()
save.image()

topoR <- bind_rows(topoR, multiweb::calc_topological_roles(noBird,nsim=100,ncores = 48) %>% mutate(Network="Nobirds"))
save.image()

mdd <- cluster_spinglass(wedd)
mdf <- tibble( modulos = list(mdd), Network="Weddell")

hub_conn <- classify_topological_roles(filter(topoR,Network=="Weddell"),wedd,mdd,TRUE) %>% mutate(Network="Weddell")

knitr::kable(aves%>% rename(name=query) %>% inner_join( hub_conn, by = "name") %>% select(Network, name:among_module_conn))

mdd <- cluster_spinglass(noBird)
mdf <- bind_rows(mdf, tibble( modulos = list(mdd), Network="Nobirds"))

hub_conn <- bind_rows(hub_conn,
classify_topological_roles(filter(topoR,Network=="Nobirds"),noBird,mdd,TRUE) %>% mutate(Network="Nobirds"))

# Module hubs and Hub connectors 
#
hub_conn %>% filter(type %in% c("modhub","hubcon")) %>% arrange(type)

# png("Figures/TopoRolNoBirds.png",width=6,height=6,units = "in",res=300)
classify_topological_roles(filter(topoR,Network=="Nobirds"),noBird,plt=TRUE)
# dev.off()
# 
# 
# mdd <- mdf$modulos[1][[1]]
# 
# png("Figures/TopoRolWeddell.png",width=6,height=6,units = "in",res=300)
classify_topological_roles(filter(topoR,Network=="Weddell"),wedd,plt=TRUE)
# dev.off()

```

## Summary of birds metrics

```{r birdsMetrics, eval=T,echo=T,message=T,warning=T}
# degree of birds joined with topoological roles 
#
V(noBird)[ degree(noBird) == 0]
siBird <- V(wedd)$name %in% aves$query

bsp <- tibble::rownames_to_column(as.data.frame(degree(wedd, v=siBird, mode=c("out")))) 
names(bsp) <- c("Species","out_degree")
bsp1 <- tibble::rownames_to_column(as.data.frame(degree(wedd, v=siBird, mode=c("in")))) 
names(bsp1) <- c("Species","in_degree")
bsp <- inner_join(bsp1,bsp) %>% arrange(desc(in_degree))
knitr::kable(bsp, digits = 3 )

bsp <- inner_join( bsp, hub_conn, by=c("Species" = "name") )

require(NetIndices)
TL <- tibble::rownames_to_column(as.data.frame(TrophInd(get.adjacency(wedd,sparse=F))))
names(TL)[1] <- "Species"

bsp <- inner_join( bsp, TL, by=c("Species") ) %>% dplyr::select(-Network,-node)

knitr::kable(bsp)
```

## Plots of the

```{r birdsMetrics, eval=T,echo=T,message=T,warning=T}

require(multiweb)
require(igraph)
require(ggraph)
require(NetIndices)

g2 <- wedd
V(g2)$name <- vegan::make.cepnames(V(g2)$name)
plot_troph_level(g2,modules =TRUE ,vertexLabel = FALSE,maxTL=5,vertex.label.color="black")

#
# Ver conexiones de aves
#
siBird <- V(wedd)$name %in% aves$query
adjacent_vertices(wedd, V(wedd)[siBird], mode="in")

adjacent_vertices(wedd, V(wedd)[c(aves$query[12],aves$query[2])], mode="in")
adjacent_vertices(wedd, V(wedd)["Electrona antarctica"], mode="in")
adjacent_vertices(wedd, V(wedd)["Euphausia superba"], mode="in")

#
# Estimar sub red de aves marinas
#
sg  <- make_ego_graph(wedd, nodes = V(wedd)[c(aves$query[12])], order=2, mode="in")

res <- do.call(igraph::union, sg)
#V(res)$name <- vegan::make.cepnames(V(res)$name)
plot_troph_level(res, vertexLabel = TRUE, tk=TRUE, vertexSizeFactor = 5,modules = TRUE, edge.width = 1)

require(NetIndices)
TL <- tibble::rownames_to_column(as.data.frame(TrophInd(get.adjacency(res,sparse=F))))
TL <- arrange(TL, desc(TL))
V(wedd)$color <- round(TL$TL)
V(wedd)$size <-  as.numeric(cut(degree(wedd),11))

```