### 2020-12-16 ModellingWeddellSea.Rmd

---
title: "Modelling Weddell Sea Food Web"
author: "L.A.S."
output: html_document
editor_options: 
  chunk_output_type: console
---

## Read Weddell Sea Data from GATEWAY

```{r readWeddell, eval=F,echo=F,message=T,warning=T}

#
# Read and check data available on GATEWAy about Weddell Sea
#
source("R/network_fun.r")
require(igraph)
require(dplyr)
require(multiweb)
require(readr)

# Read the original database
#
ga <- readr::read_csv("Data/283_2_FoodWebDataBase_2018_12_10.csv",col_types = "icccccccccccccddddddccccccccddddddcddccddciicc")
names(ga)
wedd_df <-  ga %>%filter(grepl('Weddell',foodweb.name)) # %>% dplyr::select(con.taxonomy,con.taxonomy.level,res.taxonomy,res.taxonomy.level)
#saveRDS(wedd_df, "Data/wedd_df.rds")

# Read the network file (from the same source but different format)
#
wedd <- readNetwork("Data/Wedd.dat",edgeListFormat = 2)

# Check all species are in res.taxonomy 
names(wedd_df)

# Extract the birds from the network
#
require(taxize)
can_names <- gnr_resolve(unique(wedd_df$con.taxonomy), best_match_only = TRUE, canonical=TRUE)
class_names <- tax_name(query =can_names$matched_name2  , get = "class", db = "ncbi")
aves <- class_names %>% filter( class =="Aves")

wedd_df %>% filter(con.taxonomy=="Fulmarus glacialoides") %>% select(con.mass.mean.g.)
save.image()
```

## Add body masses in g to a data.frame

```{r makeBodyMass, eval=F,echo=F,message=T,warning=T}

# Parameters for alometric equations
#
n0 <- -1.16 
g <- -0.75

# Make the body mass of Phytodetritus Sediment = 1e-14 to obtain better interaction rates for detritivores
#
wedd_df <- wedd_df %>% mutate(res.mass.mean.g.=ifelse(res.mass.mean.g.==-9.99e+02, 1e-14, res.mass.mean.g.))

res <- wedd_df %>% distinct(res.taxonomy, .keep_all = TRUE) %>% rename(name=res.taxonomy,mass_mean = res.mass.mean.g.) %>% select(name, mass_mean) 
con <- wedd_df %>% distinct(con.taxonomy, .keep_all = TRUE) %>% rename(name=con.taxonomy) %>% select(name, con.mass.mean.g.) 
con <- anti_join(con,res) %>% rename(mass_mean =con.mass.mean.g.)
wedd_sp <- bind_rows(res,con)

# Neq Multiplied by 1e8 to make min populations on the order of 1000, detritus set at big values
#
# Make the mass_mean of  Phytodetritus Sediment = 3e+07
#
wedd_sp <- wedd_sp %>% mutate(mass_mean=ifelse(mass_mean==1e-14,3e+07,mass_mean	 ),
  Neq = 10^(n0+g+rnorm(nrow(wedd_sp),0,0.2))*mass_mean^(g+1)*1e9) %>% arrange(desc(Neq)) %>% mutate(Num_eq=Neq/mass_mean)

require(ggplot2)
ggplot(wedd_sp, aes(Num_eq)) +  geom_histogram(bins=50,color="darkblue",fill="white") + theme_bw() + scale_y_log10()

require(ggplot2)
ggplot(wedd_sp, aes(Neq)) +  geom_histogram(bins=50,color="darkblue",fill="white") + theme_bw() + scale_y_log10()

# Generate interaction rates from wedd_df 
#
#
wedd_aij <- wedd_df %>% mutate(aij =  10^-3.5 * con.mass.mean.g. *( (res.mass.mean.g./con.mass.mean.g.)^0.46 / (1 + (res.mass.mean.g./con.mass.mean.g.)^2)) ) %>% select( res.taxonomy, con.taxonomy, aij) %>% arrange(desc(aij))

ggplot(wedd_aij, aes(aij)) +  geom_density(color="darkblue",fill="white") + theme_bw()

rm(res,con,can_names,class_names)
save.image()

wedd_aij %>% filter(con.taxonomy=="Fulmarus glacialoides") 
wedd_aij %>% filter(res.taxonomy=="Fulmarus glacialoides")

```