### 2022-02-02 ModellingWeddellSea.Rmd
Updated function

---
title: "Modelling Weddell Sea Food Web"
author: "L.A.S."
output: html_document
editor_options: 
  chunk_output_type: console
---

## Read Weddell Sea Data from GATEWAY

```{r readWeddell, eval=F,echo=F,message=T,warning=T}

#
# Read and check data available on GATEWAy about Weddell Sea
#
source("R/network_fun.r")
require(igraph)
require(multiweb)
require(tidyverse)

# Read the original database
#
ga <- readr::read_csv("Data/283_2_FoodWebDataBase_2018_12_10.csv",col_types = "icccccccccccccddddddccccccccddddddcddccddciicc")
names(ga)
wedd_df <-  ga %>%filter(grepl('Weddell',foodweb.name))  #saveRDS(wedd_df, "Data/wedd_df.rds")

# Read the network file (from the same source but different format)
#
wedd <- readNetwork("Data/Wedd.dat",edgeListFormat = 2)

# Check all species are in res.taxonomy 
names(wedd_df)

# Extract the birds from the network
#
# require(taxize)
# can_names <- gnr_resolve(unique(wedd_df$con.taxonomy), best_match_only = TRUE, canonical=TRUE)
# class_names <- tax_name(query =can_names$matched_name2  , get = "class", db = "ncbi")
# aves <- class_names %>% filter( class =="Aves")
# 
# wedd_df %>% filter(con.taxonomy=="Fulmarus glacialoides") %>% select(con.mass.mean.g.)

# Rename and Select fields

wedd_df <- wedd_df %>% rename(con_mass_mean=con.mass.mean.g.,res_mass_mean=res.mass.mean.g.,con_taxonomy=con.taxonomy,res_taxonomy=res.taxonomy,interaction_dim=interaction.dimensionality) %>% select(con_taxonomy,res_taxonomy, con_mass_mean,res_mass_mean,interaction_dim)

save.image()
```

## Add body masses in g to a data.frame

```{r makeBodyMass, eval=F,echo=F,message=T,warning=T}

# Parameters for alometric equations @Emmerson2004 y ... 
# @Yang2019 --> Parametriza Lotka Volterra
# @Donohue2017 --> Parametriza model derived from Yodzis and Innes (1992), updated with more recent allometric coefficients
# @Fahimipour2014 ---> Estima el flujo a partir de la tasa metabólica y las densidad de la poblacion
# @Carrara2015  --> distintos metodos para estimar interacciones basados en LV
#
# @Petchey2008 LV Parametrization an interesting idea is that it adds Type 2 functional response to check structural robustness 
#
# @Brose2006 Parametrization of consumer–resource model (Yodzis & Innes 1992) updated with new allometric coefficients (Brown et al. 2004)
# and extended to multispecies systems (Williams & Martinez 2004b)
#
# @James2015 Uses @Emerson2004 as a method to get LV interaction strength using body masses 
#
# @Neutel2014 Uses LV parametrized from empirical data 
# @Pawar2012 2D/3D search space + metabolic theory

# Check detritus
#
detritus <- wedd_df %>% filter(grepl("tritus",res_taxonomy)) 
detritus <- wedd_df %>% filter(is.na(interaction_dim))
write_tsv(wedd_df %>% arrange(interaction_dim), "Data/Wedd_mass.dat")

wedd_int <- interaction_intensity(wedd_df,res_mass_mean,con_mass_mean,interaction_dim )

#wedd_df <- wedd_df %>% mutate(res.mass.mean.g.=ifelse(res.mass.mean.g.==-9.99e+02, 1e-14, res.mass.mean.g.))

ggplot(wedd_int, aes(qRC)) +  geom_histogram(bins=50,color="darkblue",fill="white") + theme_bw() + scale_y_log10()

ave.image()

```