### 2022-12-28 Manuscript/Manuscript.tex
Update bib

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

%% copernicus_rticles_template (flag for rticles template detection - do not remove!)

%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.

%% 2-column papers and discussion papers
\documentclass[gc, manuscript]{copernicus}

%% Journal abbreviations (please use the same for preprints and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% EGUsphere (egusphere) | This is only for EGUsphere preprints submitted without relation to an EGU journal.
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Polarforschung - Journal of the German Society for Polar Research (polf)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Safety of Nuclear Waste Disposal (sand)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

% Pandoc citation processing

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
% 
% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
\begin{document}

\title{New insights into the Weddell Sea ecosystem applying a
quantitative network approach}

\Author[1, *]{Tomás I.}{Marina}
\Author[1, *]{Leonardo A.}{Saravia}
\Author[2]{Susanne}{Kortsch}

\affil[1]{Centro Austral de Investigaciones Científicas (CADIC-CONICET),
Ushuaia, Argentina}
\affil[2]{Spatial Foodweb Ecology Group, Department of Agricultural
Sciences, University of Helsinki, Finland}
\affil[*]{These authors contributed equally to this work.}

\runningtitle{New insights into the Weddell Sea food web}

\runningauthor{Marina et al.}

\correspondence{Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A.
Saravia (arysar@gmail.com)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
Network approaches can shed light on the structure and stability of
complex marine communities. In recent years, such approaches have been
successfully applied to study polar ecosystems, improving our knowledge
on how they might respond to ongoing environmental changes. The Weddell
Sea is one of the most studied marine ecosystems outside the Antarctic
Peninsula in the Southern Ocean. Yet, few studies consider the known
complexity of the Weddell Sea food web which in its current form
comprises 490 species and 16041 predator-prey interactions. Here we
analysed the Weddell Sea food web, focusing on trophic interactions that
underpin ecosystem structure and stability. We estimated the strength
for each interaction in the food web, characterised species position in
the food web using unweighted and weighted properties, and analysed
species' roles with respect to the stability of the food web. On one
hand, we found that the distribution of the interaction strength at the
food web level is asymmetric, where weak interactions are prevalent.
Including such information as a (weighted) property for species we
detected a positive relationship between species mean interaction
strength and two unweighted properties, trophic level and the total
number of interactions. We also found that only a few species are key in
terms of food web stability, presenting high mean interaction strength,
mid to high trophic level, relatively high number of interactions, and
mid to low trophic similarity. In the same analysis we have integrated
food web and species information, enabling a more complete assessment of
the ecosystem structure and function, likely highlighting the ecological
processes at play in the Weddell Sea. We consider that our results
provide new insights important for the development of effective policies
and management strategies, particularly given the ongoing initiative to
implement a Marine Protected Area (MPA) in the Weddell Sea.
\end{abstract}

\section{Introduction}

Food web analysis constitutes an important framework for understanding
ecological community structure and for conserving biodiversity through
ecosystem management \citep{Thompson2012}. Although topological food web
analysis, which considers only the presence and absence of predator-prey
interactions, provides important insights into the structure and
functioning of ecological communities
\citep[e.g.][]{Pascual2006, Kortsch2015, Marina2018, Cordone2020, Rodriguez2022},
more information on the nature of the trophic interactions is needed to
effectively charaterise ecosystem dynamics and stability
\citep[\citet{otherRefs}]{Kortsch2021}. This is a fundamental step for
providing assemessments on ecosystem vulnerability to environmental
pressures and for prioritising management actions. In this regard,
quantifying the strength of trohic interactions and species' roles
within the network are of paramount importance
\citep{Carrara2015, Allesina2015, Nilsson2016, Cirtwill2018a}.

We should maybe write a paragraph about weighted and unweighted
approaches here, where we aslo introduce the approach we used\ldots{}

\begin{quote}
{[}susanne{]} I moved the paragraph below. it makes more sense to first
talk about polar regions in genaral and then introduce the Weddell Sea.

Using a network approach, different types (e.g., terrestial, lake, etc.)
of food webs from various geographic locations are studied worldwide,
including marine polar food webs
\citep{Carscallen2012, Santana2013, Kortsch2019, Pecuchet2022}. Some of
these studies from the Arctic show how food web properties (e.g.,
connectance) are constrained by environmental factors such as sea ice
cover and seawater temperature \citep{Kortsch2019, Pecuchet2022}. In the
Southern Ocean, important insights have been gained into mechanisms of
energy flow, the relative importance and traits of individual species
and the influence of environmental variables (e.g., sea-ice) on the
structure of local food webs \citep{Cordone2020, Rossi2019}.
{[}susanne{]} what do you mean here: the relative importance and traits
of individual species? What important mechanisms? Can we be more
specefic? Can we also be more specific about how environmental variables
(e.g., sea-ice) influence the structure of local food webs in the SO?
\end{quote}

The Weddell Sea is expected to be one of the last regions of the
Southern Ocean to experience the consequences of climate change due to
its extensive ice cover and ocean currents \citep{Teschke2021} resulting
in less sea surface warming compared to other areas of the Southern
Ocean. Because of these environmental characteristics, the Weddell Sea
may serve as a refuge for Antarctic species which depend on sea ice
(e.g.~krill, emperor penguin, Weddell seal) or have low heat tolerance
(e.g., most notothenioid fishes) due to their adaptations to freezing
tempeartures \citep{Griffiths2017}. While essential large-scale
hydrodynamic relationships are relatively well-known for this region
\citep{deSteur2019}, information on the current distribution, abundance
and sensitivity to climate change is only partially known for a few
species (e.g., XXXXX) \citep{Houstin2022}. \textgreater{}
{[}susanne{]}it is not clear from the text why Weddell Sea is one of the
last regions of the SO to face the consequences of warming, be explicit
here, how are the currents affecting the Weddell Sea? \textgreater{}

Overall, the complexity the Weddell Sea food web is high and consists of
488 species and 16200 predator-prey interactions\citep{Jacob2011}. In an
attempt to better understand species role related to food web stability,
Jacob et al.~(2011) performed secondary extinction experiments and found
that the removal of small to medium-sized, and not large, organisms
caused a cascade of secondary extinctions. This findings highlighted the
relative importance of predators, rather than prey, in the architecture,
functioning and stability of the Weddell Sea food web, which coincides
with findings from recent meta-analyses in natural complex food webs
\citep{Brose2019, Perkins2022}. Other investigations have considered
this food web though in a meta-analysis context\ldots.{[}more text here,
WHAT DO THESE META\_ANALSYES STUDIES SHOW{]}
\citep{Brose2019, Perkins2022, Saravia2022}.

Several methodologies have been applied to estimate the interaction
strength in food webs, where the quantity and quality of data mostly
determine which approach is more convenient \citep{Berlow2004}.
Experimental methods combined with dynamic modelling
\citep{Emmerson2004, Carrara2015}, measurements of species abundances
trough time \citep{Fahimipour2014, Chang2021}, and/or estimation of
metabolic rates and biomass of all species in the community
\citep{Neutel2014}; all these methods require a large experimental
and/or sampling effort that restrict them to simple networks (circa 10
species). Other methods based on predator-prey body mass allometric
relationships \citep{Berlow2009, Pawar2012} could be applied to larger
networks due to less data requirements, although they might have less
power with increasingly complex networks unless more traits are included
in the estimations \citep{Jonsson2018}. For food webs composed by nearly
1000 species and about 20000 interactions only those methods with less
data requirements are feasible. On of these is the one proposed by
\citet{Pawar2012}, which allows to obtain interaction strength
estimations by combining data on body mass (of consumer -predator- and
resource -prey-) and interaction dimensionality (classified as 2 or 3
dimensions).

In this study, we aim to go beyond a purely topological
(presence/absence) assessment of who eats whom in the Weddell Sea
ecosystem by providing a quantitative analysis of the trophic
interaction network. We aim to analyse role for the structure and
stability of the food web. To achieve this, we: 1) estimated the
strength for each interaction in the Weddell Sea food web, 2)
characterised species role considering both weighted and unweighted
properties, and 3) analysed the species' role realted to the stability
of the food web. This is the first time that the interaction strength
was estimated for all the interactions mostly at the species level for
the Weddell sea food web

\section{Methodology}

\subsection{Study area}

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS,
stretching approximately 450 km in length from East to West (Figure 1).
Water depth varies between 200 and 500 meters, and shallower areas are
covered by continental ice, which forms the coastline along the eastern
and southern part of the Weddell Sea. The shelf area contains a complex
three-dimensional benthic habitat with large benthic biomasses,
intermediate to high diversity in comparison to benthic boreal
communities and a spatially patchy distribution of organisms
\citep{Dayton1990, Teixido2002}.

\subsection{Weddell Sea food web dataset}

The Weddell Sea food web was retrieved from the GlobAL daTabasE of
traits and food Web Architecture (GATEWAy, version 1.0) of the German
Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
\citep{Brose2018}. In addition to predator-prey interactions, the
database contains information on other biological data such as the mean
body mass and movement type for each species in the food web.
Furthermore, it incorporates information about the interaction itself,
such as the dimension of the predator search space (2 or 3 dimensions).
In its current form, the Weddell Sea food web comprises 490 species and
16041 predator-prey interactions and constitutes one of the most
resolved food webs constructed to date \citep{Jacob2011}.

\subsection{Dataset analyses}

\subsubsection{Interaction strength estimation and distribution}

To estimate the strength of each pairwise interaction in the food web,
we followed an approach proposed by \citet{Pawar2012}. The minimum data
requirements are body mass of the consumer (predator) and resource
(prey), and the interaction dimensionality (ID) classified as 2 or 3
dimensions. The ID is defined as the search space dimensions of the
predator, which is also equivalent to the movement space of the prey.
Thus, the ID is classified as 2D when both predator and prey move in 2D
(e.g., both are benthic) or if a predator moves in 3D and a prey in 2D
(e.g., pelagic predator on benthic prey). The ID is classified as 3D
when both predator and prey move in 3D (e.g., both pelagic) or if the
predator moves in 2D and the prey in 3D (e.g., benthic predator, pelagic
prey) \citep{Pawar2012}. GATEWAy v.1.0 provides information on the mean
body mass for consumers and resources, except for `detritus' and
`sediment', and the dimensionality for the majority of the interactions,
though the latter is missing in some cases (924 interactions). To
complete the missing data on species `dimensionality', we used
information about the movement type of predators and prey included in
GATEWAy.

The main equation we used for estimating the interaction strength IS
was:

\begin{equation}
IS = \alpha x_R \frac{m_R}{m_C}
\end{equation}

where \vec{\alpha} is the search rate, \vec{x_R} is the resource
density, and \vec{m_R} and \vec{m_C} are the body mass for the resource
and the consumer, respectively \citep{Pawar2012}.

We obtained estimates for resource density and the search rate from the
scaling relationships with the resource and the consumer mass,
respectively \citep{Pawar2012}. The coefficients of such relationships,
determined by ordinary least squares regression, vary with the
interaction dimensionality. On one hand, resource density scales with
resource mass as power-law with different exponents in 2D and in 3D.
Since mean mass for resources `phytodetritus' and `sediment' were not
available in GATEWAy, we considered the body mass of the smallest
phytoplankton species (`Fragilariopsis cylindrus') as a proxy. This is
justified by the fact that `phytodetritus' and `sediment' are mainly
composed of dead or senescent phytoplankton reaching the seabed
\citep{Wolanski2011}. On the other hand, search rate scales with
consumer mass as power-law with exponents in 2D and in 3D.

We fitted six candidate models (Exponential, Gamma, log-Normal, Normal,
Power-law and Uniform) the interaction strength distribution using
maximum likelihood \citep{McCallum2008}, and selected the best fitting
model by computing the Akaike Information Criterion AIC
\citep{Burnham2002}.

\subsubsection{Species properties}

To characterise the role of each species in the food web, we considered
unweighted and weighted food web properties (Figure 2). Unweighted
properties are related to properties commonly used in qualitative food
web studies and only describe the presence or absence of interactions
without any information on strength between a pairwise species link
\citep{Martinez1991, Dunne2002, Borrelli2014}. In contrast, weighted
properties capture the importance of a trophic interaction by
considering its strength.

To assess species roles as a function of the weighted food web, we
focused on mean interaction strength defined as the average strength of
all interactions for a given species. Further we calculated three
unweighted species properties: a) species degree, i.e., the sum of in-
and out-going interactions ; b) trophic level ; and c) trophic
similarity, i.e., the trophic overlap based on shared and unique
resources and consumers. These metrics were chosen to assess a species
role based on the unweighted network. The species degree has often been
equated with species importance to the structure and functioning within
a food web, i.e.~perturbations to high-degree species may therefore have
more significant effects on the food web robustness to perturbations
than low-degree species
\citetext{\citealp{Dunne2002a}; \citealp[references
in][]{Cirtwill2018a}}. The trophic level offers information about how
important a species is to its biotic community, i.e., top predators and
primary producers are expected to have particularly large effects on the
rest of their communities through top-down and bottom-up control,
respectively \citep[references in][]{Cirtwill2018a}. Trophic similarity
is an index of trophic overlap considering the set of prey and predators
for a pair of species; it measures one of the most important aspects of
species' niches, the trophic niche, and functional aspects of
biodiversity \citep{Martinez1991, Williams2000}.

Furthermore, we took a species' habitat into account, which describes
the physical position of a species within the ecosystem. Species were
categorised as: 1) benthic, if a species lives on the seafloor; 2)
pelagic, if a species lives close to the surface; 3) benthopelagic, if
it moves between and connects the mentioned environments; 4) demersal,
if it lives and feeds on or near the bottom of the sea; and 5)
land-based, if the consumer is not strictly aquatic but feeds
predominantly on marine species. Species habitat affiliation was
retrieved from \citet{Jacob2011}.

With the aim of studying the relationship between the interaction
strength of the species (weighted property) and its unweighted
properties we performed linear regression analyses between the log mean
interaction strength and each of the mentioned unweighted properties.
Thus we considered the interaction strength as the dependent variable
and the given unweighted property as the independent variable, and
obtained the coefficients (slope and intercept) for the linear model.
Models were fitted using the least squares approach. We also explored
the mean interaction strength distribution with the species habitat.

Formulas used to obtain the above species properties are described in
Supplementary Material.

\subsubsection{Extinction simulations and stability}

To analyse the impact of species on food web stability, we performed
extinction simulations deleting one species at a time, that is for every
extinction network size was reduced by one species only. After each
extinction, we calculated the stability of the network minus the
targeted species (489 nodes) and compared it with that of the whole
network (490 nodes in total). To calculate stability, we used the mean
of the real part of the maximum eigenvalue of the Jacobian matrix using
randomized Jacobians and keeping the predator-prey sign structure fixed
\citep{Allesina2008, Grilli2016}. This index indicates a more stable
food web when it is negative. We performed 1000 simulations for each
species extinction and obtained a mean maximum eigenvalue for each case.
At last we statistically analysed such a difference with an
Anderson-Darling test considering a p-value \textless{} 0.01
\citep{Scholz1987}. If this difference is positive, then the stability
of the food web is higher if a species was removed, and vice versa. A
detailed description on the stability calculations can be found in
Supplementary Material.

Once we had the results for the impact on stability for each species
extinction, we plotted them considering weighted (interaction strength)
and unweighted properties, and species habitat. With this we aim to
characterise those species with a relatively high effect on the
stability of the food web.

All analyses were performed in R software, using the R packages igraph
\citep{Csardi2005}, cheddar \citep{Hudson2013}, and multiweb
\citep{Saravia2019}. The source code and data are available at
https://github.com/EcoComplex/WeddellSea.

\section{Results}

\subsection{Interaction strength distribution}

The statistical distribution that best fitted the empirical interaction
strength distribution of the Weddell Sea food web was a `gamma' due to
the high proportion of weak interactions and the existence of a few
strong interactions (Figure 3, Table S3).

\subsection{Species' role related to their mean interaction strength}

We found that the species' mean interaction strength (weighted property)
shows different relationships with the unweighted properties analysed
(Figure 4A-D). In this regard, there is a positive relationship between
interaction strength and trophic level, i.e., the higher the trophic
level of the species, the higher its mean interaction strength. We also
found a significant but less evident positive relationship with species
degree. Contrary, there was no significant relationship between mean
interaction strength and trophic similarity. Considering species habitat
affiliation, the ``Benthopelagic'' and ``Pelagic'' categories contained
the two species with the highest mean interaction strength, the killer
whale Orcinus orca and the colossal squid Mesonychoteuthis hamiltoni,
respectively. However, the majority of the species with relatively
higher interaction strength belonged to the ``Demersal'' and
``Land-based'' habitats groups. Species inhabiting the benthic realm
showed the lowest mean interaction strength (Figure 4D).

\subsection{Species impact on food web stability}

Our extinction analyses showed that the majority of species had no
significant impact on food web stability after being removed (Figure 5).
Most of the species (black points in figure 5) did not change the
stability of the network considerably after being removed, except for a
few species. Only 15 out of 490 species (3.06\%) gave rise to
significant changes in the food web's stability after their removal
(Table 2). Most of these species had a positive impact on food web
stability, i.e., network stability increased after their removal. Only
two species significantly decreased network stability after being
removed, the demersal fish Pagetopsis macropterus and the benthopelagic
amphipod Maxilliphimedia longipes.

After exploring the stability difference against the species properties
(Figure 5), we found that the species that generated a significant
impact on the stability of the food web were characterised by: 1) high
mean interaction strength; 2) mid to high trophic levels (TL
\textgreater{} 3.2); 3) relatively high number of interactions (Degree
\textgreater{} 25); and 4) mid to low trophic similarity (TS \textless{}
0.16). Habitat wise, species with a significant impact on the stability
were present in all habitats, except for the benthic realm. Table 2
shows the results for the species with highest impact on the food web
stability.

\section{Discussion}

\subsection{Many weak and a few strong interactions}

Our analyses show that the distribution of species interaction strength
at the network level is asymmetric, i.e., the Weddell Sea food web
contains many weak interactions and only a few strong ones. This finding
is consistent with many previous theoretical and empirical studies
\citep[e.g.][]{McCann1998, Neutel2002, Emmerson2004, Wootton2005, Kortsch2021}.
The asymmetric distribution of interaction strength in food webs has
been interpreted as an explanation for the persistence of complex
communities in nature \citep{Bascompte2005, Allesina2015, Nilsson2016}.
Here we show that this pattern is also prevalent in the Weddell Sea, one
of the most complex food webs to date, comprising 490 species and 16041
predator-prey interactions. This finding is in someway validating the
method we used, validation of allometric methods of IS estimation not
including interaction dimensionality has been performed for microcosmos
of relatively few species \citep{Jonsson2018}.

\subsection{Species's role related to their mean interaction strength}

We employed a range of descriptors using both unweighted and weighted
food web properties to characterise the dynamic and multifaceted nature
of the Weddell Sea food web. Our results show a positive relationship
between interaction strength and trophic level, and between interaction
strength and species degree. Mean interaction strength increases with
trophic level and species degree. The former relationship contradicts
studies that suggest that mid-trophic level species (which, krill???
examples are always good) are involved in the major pathways of energy
flow in high-latitude marine ecosystems
\citep{Pinkerton2014, Murphy2016, McCormack2020, Riccialdelli2020}.
However, this finding could also be explained by the lack of species
biomass information in the calculation of interaction strength using the
methodology we applied \citep{Parwar2012}. The methodlogy by Parwar et
al.(2012) allows information on species density/biomass to be included,
but this type of data was not available for the majority of species in
our food web. In the Weddell Sea, species with high degree also tend to
have high mean interaction strengthsThe positive relationship between
interaction strength and species degree reinforces the central role of
species with many interactions: species with a high degree (hubs) have a
large impact on overall food web structure and functioning
\citep{Dunne2002a, Kortsch2015}. This information on the quantity and
quality of interactions and its relationship enables a robust assessment
of the species' role in the stability of the food web
\citep{Cirtwill2018a}.

\subsection{Species impact on food web stability}

Only a few species play a key role with respect to the Weddell Sea food
web stability, according to the mean maximum eigenvalue stability index
employed in this study. This is in concordance with other studies on
complex empirical food webs in marine ecosystems in the Arctic and other
locations in Antarctica \citep{Kortsch2015, Marina2018, Rodriguez2022}.
These key species are characterised by a particular set of food web
properties: high to mean IS; mid to high trophic level; a relatively
high number of interactions; and mid to low trophic similarity. In a
previous study on sequential extinction simulations for the Weddell Sea
food web \citep[!!]{CITA}, it was found that larger bodied-sized species
could be lost without causing a collapse of the network. A major caveat
of this finding, also recognised by the authors, was that population
dynamics were ignored and hence no top-down extinctions, or other
indirect effects, could occur. In our study we considered such top-down
effects by including information on the species IS, which is of
paramount importance when analysing the response of perturbations in
ecological communities \citep{McCann1998, Montoya2009, Novak2011}. Thus,
our study suggests that species with high mean IS and high trophic
levels need to be considered with particular attention when trying to
predict the effects of perturbations on the Weddell Sea ecosystem. This
conclusion is further reinforced by the finding that these species have
mid to low trophic similarity, which means that few other species of the
food web can occupy the same trophic role. In a review, it was
emphasised that polar pelagic communities are particularly sensitive to
changes due to a low functional redundancy at key trophic levels
\citep{Murphy2016}. Here we provide broader analyses of species impact
on food web robustness by including species from all habitats (benthic,
pelagic and land-based). This suggests that the sensitivity of marine
polar ecosystems to environmental perturbations is a concern also beyond
the pelagic realm.

\clearpage
\conclusions[Conclusions]

Our study goes beyond the current understanding of how species influence
ecosystem structure and stability in the Weddell Sea in particular and
in most polar regions in general \citep{Murphy2016, McCormack2021}. In
the same analysis we have integrated information about weighted
(interaction strength) and unweighted species properties, enabling a
more complete assessment of the species' role in the food web structure
and function.

We consider that the information provided in this study is important for
the development of effective policies and management strategies,
particularly given the ongoing initiative to implement a Marine
Protected Area (MPA) in the Weddell Sea \citep{Teschke2021}.

\clearpage

\begin{figure}
\includegraphics[width=12cm]{Fig.1_StudyMap} \caption{Map of the Weddell Sea and Dronning Maud Land sector highlighting the high Antarctic shelf as a dashed-line contour. Modified from www.soos.aq.}\label{fig:unnamed-chunk-1}
\end{figure}

\clearpage

\begin{figure}
\includegraphics[width=12cm]{Fig.2_ToyFoodWeb} \caption{Scheme of a network showing the weighted and unweighted properties we used to characterize the species of the Weddell Sea food web.}\label{fig:unnamed-chunk-2}
\end{figure}

\clearpage

\begin{figure}
\includegraphics[width=12cm]{Fig3_IntDist} \caption{Frequency distribution of interaction strengths for the Weddell Sea food web. Total number of interactions = 16041. The distribution was best fitted to a ‘gamma’ model.}\label{fig:unnamed-chunk-3}
\end{figure}

\clearpage

\begin{figure}
\includegraphics[width=12cm]{Fig4_LinReg} \caption{Relationships between weighted (mean Interaction Strength) and unweighted properties including habitat. Linear regressions are shown between log(mean interaction strength) and trophic level (A), degree (B) and trophic similarity (C). Linear regressions for trophic level ($y = 1.12x - 15.29, R^2 = 0.43, p-value < 2e-16$), degree ($y = 0.006x - 12.77, R^2 = 0.03, p-value = 4.06e-5$) and trophic similarity ($y = -1.46x - 12.18, R^2 = -0.0004, p-value = 0.36$).}\label{fig:unnamed-chunk-4}
\end{figure}

\clearpage

\begin{figure}
\includegraphics[width=12cm]{Fig.5_QSSDif} \caption{Stability  difference (mean maximum eingenvalue) between the whole Weddell Sea food web (n = 490) and the food web minus one species (n = 489) for weighted (interaction strength) and unweighted species properties, and habitat. Point color indicates the impact on the stability; if significant the extinction of that species altered the stability of the food web.}\label{fig:unnamed-chunk-5}
\end{figure}

\clearpage

\begin{table}[t]
\caption{Properties of the species that when become extinct generated a significant impact on the stability of the Weddell Sea food web, ordered by significance (Anderson-Darling p-value). References: meanIS = mean interaction strength, TL = trophic level, Deg = degree, TS = trophic similarity, StabDif = stability difference, ADvalue = Anderson-Darling p-value.}
\begin{tabular}{l c c c c c c c}
\tophline

\textbf{Species} & \textbf{meanIS} & \textbf{TL} & \textbf{Deg} & \textbf{TS} & \textbf{Habitat} & \textbf{StabDif} & \textbf{ADvalue}\\
\middlehline
Orcinus orca & 1.83e-4 & 5.03 & 26 & 0.037 & Benthopelagic & 4.67e-5 & 2.28e-41 \\
\middlehline
Macrourus holotrachys & 8.30e-5 & 4.70 & 85 & 0.112 & Benthopelagic & 3.55e-5 & 2.73e-23 \\
\middlehline
Pagetopsis macropterus & 7.08e-5 & 4.64 & 76 & 0.113 & Demersal & -1.80e-5 & 2.38e-12 \\
\middlehline
Abyssorchomene nodimanus & 2.56e-5 & 4.21 & 137 & 0.130 & Benthopelagic & 2.30e-5 & 8.52e-10 \\
\middlehline
Dissostichus mawsoni & 7.82e-5 & 4.12 & 87 & 0.126 & Pelagic & 2.17e-5 & 1.57e-9 \\
\middlehline
Macrourus whitsoni & 7.14e-5 & 4.55 & 92 & 0.124 & Benthopelagic & 2.12e-5 & 3.30e-8 \\
\middlehline
Hydrurga leptonyx & 1.03e-4 & 4.72 & 67 & 0.094 & Land-based & 2.04e-5 & 9.66e-6 \\
\middlehline
Mesonychoteuthis hamiltoni & 1.80e-4 & 4.41 & 29 & 0.028 & Pelagic & 1.82e-5 & 4.59e-5 \\
\middlehline
Champsocephalus gunnari & 7.62e-5 & 3.72 & 46 & 0.086 & Pelagic & 1.83e-5 & 6.79e-5 \\
\middlehline
Notothenia marmorata & 8.27e-5 & 4.09 & 44 & 0.091 & Demersal & 1.60e-5 & 1.23e-4 \\
\middlehline
Arctocephalus gazella & 9.28e-5 & 4.67 & 61 & 0.093 & Land-based & 1.17e-5 & 2.09e-4 \\
\middlehline
Trematomus pennellii & 3.04e-5 & 4.04 & 192 & 0.158 & Demersal & 1.44e-5 & 1.00e-3 \\
\middlehline
Mirounga leonina & 1.20e-4 & 4.87 & 56 & 0.080 & Land-based & 1.41e-5 & 1.28e-3 \\
\middlehline
Notothenia coriiceps & 4.94e-5 & 4.27 & 130 & 0.126 & Demersal & 1.44e-5 & 1.66e-3 \\
\middlehline
Maxilliphimedia longipes & 2.21e-6 & 3.26 & 60 & 0.136 & Benthopelagic & -4.46e-6 & 9.74e-3 \\

\bottomhline
\end{tabular}
\end{table}

\appendixfigures
\clearpage

\begin{figure}
\includegraphics[width=12cm]{App1_FWplot} \caption{Graphic representation of the Weddell Sea food web. Species (nodes) are arranged vertically and colored by trophic level. The diameter of the node indicates the total number of interactions. Predator-prey interactions are represented by the arrows, from prey to predator.}\label{fig:unnamed-chunk-6}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontribution{TIM and LAS: Conceptualization (lead); Data curation
(lead); Formal analysis (lead); Methodology (lead); Coding (lead);
Writing -- original draft (lead); Writing -- review and editing (lead).
SK: Conceptualization (lead); Formal analysis (supporting); Methodology
(supporting); Coding (supporting); Writing -- original draft
(supporting); Writing -- review and editing
(supporting).} %% optional section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\competinginterests{The authors declare no competing
interests.} %% this section is mandatory even if you declare that no competing interests are present

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
Thanks to the rticles contributors!
\end{acknowledgements}

%% REFERENCES
%% DN: pre-configured to BibTeX for rticles

%% The reference list is compiled as follows:
%%
%% \begin{thebibliography}{}
%%
%% \bibitem[AUTHOR(YEAR)]{LABEL1}
%% REFERENCE 1
%%
%% \bibitem[AUTHOR(YEAR)]{LABEL2}
%% REFERENCE 2
%%
%% \end{thebibliography}

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
\bibliographystyle{copernicus}
\bibliography{../WeddellSea.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

\end{document}